Modelling last glacial cycle ice dynamics in the Alps

### 2017-08-23 articles/cycle/alps-cycle.tex
Draft discusion of maximum extent.

    In this section, we compare the model output to geological evidence from
    the last glacial cycle, in terms of maximum extent and ice flow patterns
    (Sect.~\ref{sec:flow}), timing of the Last Glacial Maximum
    (Sect.~\ref{sec:timing}), ice thickness (Sect.~\ref{sec:thickness}), and
    glacial cycle dynamics (Sect.~\ref{sec:glaciations}).

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Maximum configuration}
\label{sec:flow}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    \begin{itemize}
    \item Maximum extent is reached at 24.57 ka (Fig.~\ref{fig:lgmvel}b).
    \item Some glaciers are not at their maximum (Fig.~\ref{fig:lgmvel}a).
    \item Maximum extent is farily well reproduced (Fig.~\ref{fig:lgmvel}a).
    \item Peripheral ice caps well reproduced (Fig.~\ref{fig:lgmvel}a).
    \item Too much ice in the east (Fig.~\ref{fig:lgmvel}a).
    \item Too little ice in the west (Fig.~\ref{fig:lgmvel}a).
    \item Maximum stages flow pattern is complex (Fig.~\ref{fig:lgmvel}a).
    \item Flow goes sometimes across topography (Fig.~\ref{fig:lgmvel}a).
    \item Transfluences (Montgenèvre, col de l'Échelle, Simplon, Albrun, Gries,
                        (Brünig, Majola, Tagliamento)
    \end{itemize}

>>>>

    In this section, we compare the model output to geological evidence from
    the last glacial cycle, in terms of Last Glacial Maximum extent
    (Sect.~\ref{sec:extent}), ice flow patterns
    (Sect.~\ref{sec:flow}), timing of the Last Glacial Maximum
    (Sect.~\ref{sec:timing}), ice thickness (Sect.~\ref{sec:thickness}), and
    glacial cycle dynamics (Sect.~\ref{sec:glaciations}).

@@

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Last Glacial Maximum ice extent}
\label{sec:extent}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    The modelled total ice volume reaches a maximum of 122.5\e{3}\,\unit{km^2},
    or 300.4\,mm of sea-level equivalent, at 24.56\,ka
    (Fig.~\ref{fig:lgmvel}b). A maximum glacierized area of
    163.3\e{3}\,\unit{km^2} is attained shortly aftewards at 24.57\,ka.
    (Fig.~\ref{fig:lgmvel}b). Although the modelled timing of the LGM varies
    accross the mountain range (Sect.~\ref{sec:timing}), at 24.57\,ka nearly
    all glaciers extend to within a few km from their modelled maximum stage
    (Fig.~\ref{fig:lgmvel}a, dashed orange line).

    Although the palaeo-climate forcing was adapted (Sect.~\label{sec:climate})
    to model this maximum configuration in broad agreement with the geological
    reconstruction of the LGM extent (Fig.~\ref{fig:lgmvel}a, solid red line),
    local discrepancies remain. Most notably, reconstructed ice extent in the
    north-western Alps, including the Rhone Glacier complex, the Jura ice cap,
    and the Lyon Lobe, is largely missing from our model results. On the other
    hand, the model yields excessive glaciation in the eastern Alps, where the
    Palten, Drau and Sava glaciers are modelled to extend tens of kilometers
    beyond the mapped moraines. These discrepancies might indicate that the
    LGM climate was characterized by an east-west gradient in temperature or
    precipitation anomalies relative to present not included in the climate
    forcing, or perhaps a east-west gradient in variables not accounted for
    by the PDD model, such as cloud cover or dust deposition.

    On a more regional level, the modelled maximum extent of the Ticino, Iseo,
    Garda and Piave glaciers in the central southern Alps is constrained
    several kilometres within the mapped LGM moraines. This appears to result
    from the paleo-precipitation reduction used to force the model
    (Fig.~\ref{footprints}b), which is likely unrealistic in at least this part
    of the model domain. On the other hand, the Durance glacier in the
    south-western Alps is modelled to extend several kilometers beyond the
    mapped limit, indicating that the LGM temperature depression was likely
    dampened by Mediterranean climate in this part of the model domain.

    The model reproduces peripheral ice caps where documented by geological
    evidence on the Vercors, Chartreuse and Bauge prealpine reliefs,
    the Jura Mountains, the Vosges Mountain, the Black Forest, the Bohemian
    Forest and the Dinaric Alps. An independent ice cap also covers the
    Hochwart massif during most of the simulation, yet it is engulfed by
    alpine glaciers during the LGM to become a peripheral ice dome. The LGM
    extent of peripheral ice caps is underestimated in the Vosges and Jura
    mountains, while it is overestimated for the Vercors. There is a conformity
    between the discrepancies between modelled and reconstructed LGM extent
    for the main alpine glacier complex vs peripheral ice caps, which indicates
    a climatic cause rather than an ice dynamics cause.

    \todo{add location names on Fig.~\ref{fig:lgmvel}.}

% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -
\subsection{Ice flow patterns}
\label{sec:flow}
% -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -- -

    \begin{itemize}
    \item Maximum stages flow pattern is complex (Fig.~\ref{fig:lgmvel}a).
    \item Flow goes sometimes across topography (Fig.~\ref{fig:lgmvel}a).
    \item Transfluences (Montgenèvre, col de l'Échelle, Simplon, Albrun, Gries,
                        (Brünig, Majola, Tagliamento)
    \end{itemize}
